---
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2:
    template: template/brief_template.tex
    citation_package: biblatex
documentclass: book
bibliography: content/references.bib
---

# Tables {#tables} 

## Making LaTeX tables play nice
Dealing with tables in LaTeX can be painful.
This section explains the main tricks you need to make the pain go away.

(Note: if you are looking at the ebook version, you will not see much difference in this section, as it is only relevant for PDF output!)

### Making your table pretty
When you use `kable` to create tables, you will almost certainly want to set the option `booktabs = TRUE`.
This makes your table look a million times better:

```{r, message=FALSE}
library(knitr)
library(tidyverse)

head(mtcars) %>% 
  kable(booktabs = TRUE)
```

\vspace{4mm}

Compare this to the default style, which looks terrible:

```{r}
head(mtcars) %>% 
  kable()
```

### If your table is too wide
You might find that your table expands into the margins of the page, like the tables above.
Fix this with the `kable_styling` function from the [`kableExtra`](https://haozhu233.github.io/kableExtra/) package:

```{r, message=FALSE}
library(kableExtra)

head(mtcars) %>% 
  kable(booktabs = TRUE) %>% 
  kable_styling(latex_options = "scale_down")
```

This scales down the table to fit the page width.

### If your table is too long
If your table is too long to fit on a single page, set `longtable = TRUE` in the `kable` function to split the table across multiple pages.

```{r}
a_long_table <- rbind(mtcars, mtcars)

a_long_table %>% 
  select(1:8) %>% 
  kable(booktabs = TRUE, longtable = TRUE)
```

When you do this, you'll probably want to make the header repeat on new pages.
Do this with the `kable_styling` function from `kableExtra`:

```{r}
a_long_table %>% 
  kable(booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = "repeat_header")
```

Unfortunately, we cannot use the `scale_down` option with a `longtable`. 
So if a `longtable` is too wide, you can either manually adjust the font size, or show the table in landscape layout. 
To adjust the font size, use kableExtra's `font_size` option:

```{r}
a_long_table %>% 
  kable(booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 9, latex_options = "repeat_header")
```
